---
title: "Lab 08"
author: "Nichelle Camden"
format: html
editor: visual
toc: TRUE
---

## Unsupervised Learning Analysis of Human Breast Cancer Cells

#1. Exploratory data analysis
Importing the data
```{r}
fna.data <- "WisconsinCancer.csv"
wisc.df <- read.csv(fna.data, row.names=1) 
#head (wisc.df)
```
Assigning the data to an object called `fna.data` and saving the input file into the project directory.
Then, storing as wisc.df

Getting rid of the first column because it's unnecessary.
```{r}
wisc.data <- wisc.df[,-1]
```
Creating a new vector for the diagnosis column to use for later. 
```{r}
diagnosis <- as.factor(wisc.df$diagnosis)
```

##Q1. How many observations are in this dataset?
```{r}
dim(wisc.df)
```
There are 569 rows and 31 columns of observations in this dataset.

##Q2. How many of the observations have a malignant diagnosis?
```{r}
wisc.mal <- table(wisc.df$diagnosis)
wisc.mal
```
There are 212 malignant diagnoses.

##Q3. How many variables/features in the data are suffixed with _mean?
```{r}
grep("_mean", colnames(wisc.data))
```
There are 10 columns suffixed with "_mean"

# 2. Principal Component Analysis

To determine if the data need to be scaled, check the column means and standard deviation
```{r}
colMeans(wisc.data)
apply(wisc.data, 2, sd)
```

Execute PCA with the prcomp() function on the wisc.data, scaling if appropriate, and assign the output model to wisc.pr.
```{r}
wisc.pr <- prcomp(wisc.data)
summary(wisc.pr)
```





