---
title: "Class 5: Data Visualization with ggplot"
author: "Nichelle Camden"
format: html
toc: true
---

# Our first plot

R has base graphics

```{r}
plot(cars)
```

How would I plot this with `ggplot2`?
No, we need to install and load the ggplot2 package first.
To install any package in R, we use the `install.packages()` function, and put ggplot in quotes because it is a string.

```{r}
# install.packages("ggplot2")
```

Before I can use this package I need to load it with a `library()` call
```{r}
library(ggplot2)
#ggplot(cars) <- won't work yet, needs more layers
```

Every ggplot needs at least 3 layers:

- **Data** (the dara.frame we have),
- **Aes** (the aesthetic mapping of our data to what we want to plot)
- **Geoms** (How we want to plot this stuff)

```{r}
ggplot(data=cars) + 
  aes(x=speed, y=dist) + geom_point()
```

Add another geom

```{r}
ggplot(data=cars) + 
  aes(x=speed, y=dist) + 
  geom_point() + 
  geom_smooth(method = lm, se = FALSE)
```


## A more interesting plot

```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```
```{r}
nrow(genes)
colnames(genes)
ncol(genes)
table(genes$State)


```

There are `r nrow(genes)` genes in this data set

```{r}
ggplot(genes) +
  aes(x = Condition1, y = Condition2) +
  geom_point()
```

```{r}
p <- ggplot(genes) +
  aes(x = Condition1, y = Condition2, col = State) +
  geom_point()

p + scale_colour_manual( values=c("blue","gray","red") ) + 
  labs(title = "Gene Expression Changes Upon Drug Treatment", x= "Control (no drug)", y= "Drug Treatment" )
```

```{r} #/ eval: false
install.packages("gapminder")
library(gapminder)
```

```{r} #/ eval: false
install.packages("dplyr")
library(dplyr)

gapminder_2007 <- gapminder %>% filter(year==2007)

ggplot(gapminder_2007) +
  aes(x=gdpPercap, y=lifeExp) +
  geom_point()
```

