---
title: "Lab6_HW"
author: "Nichelle Camden"
format: html
editor: visual
---

## Lab 6 HW 

First, I'm going to install the bio3d package and call the library.

Then I'm going to call in the pdf files and name them s1-3
```{r}
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug
```

`trim.pdb` produces a smaller PDB object that contains a subset of atoms from the larger object. Here, we're grabbing chain A from each protein and naming them in correlation with the new subset.
```{r}
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")
```

Creating and naming more specific subsets
```{r}
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
```


`plotb3` is going to create a scatter plot (here with a line aesthetic) with the option to add secondary structures in the marginal regions that should correlate with the protein (ie "sse=" should match protein)
```{r}
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

```{r}
x <- s1.b
plotb3(x, sse=s1.chainA, typ="l", ylab="Bfactor")

x <- s2.b
plotb3(x, sse=s1.chainA, typ="l", ylab="Bfactor")

x <- s2.b
plotb3(x, sse=s2.chainA, typ="l", ylab="Bfactor")

#need to figure out how to make the "sse=" part interchangeable with the protein subset name so the margins are correct

```
```{r}
y <- s2.chainA

x <- s2.b
plotb3(x, sse=y, typ="l", ylab="Bfactor")
```
```{r}
x <- s1.b
y <- s1.chainA
plotb3(x, sse=y, typ="l", ylab="Bfactor")

x <- s2.b
y <- s2.chainA
plotb3(x, sse=y, typ="l", ylab="Bfactor")

x <- s3.b
y <- s3.chainA
plotb3(x, sse=y, typ="l", ylab="Bfactor")
```
now to put everything together

```{r}
visualplot <- function(x) {
  plotb3(x, sse=y, 
  typ="l", ylab="Bfactor")
}

visualplot(s1.b)
visualplot(s2.b)
visualplot(s3.b)
#these all have the same margins, so, not correct
```
I think I'm getting closer but the margins are not right :/
